---
title: "Final Project"
author: "Aaron Abraham Mathews"
date: "2025-01-16"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction
The goal of this exercise is to create a product to highlight the prediction algorithm that you have built and to provide an interface that can be accessed by others. For this project you must submit:<br>


## Algorithm used to make prediction
```{r,echo=TRUE}
predictWord <- function(the_word) {
    word_add <- stripWhitespace(removeNumbers(removePunctuation(tolower(the_word),preserve_intra_word_dashes = TRUE)))
    # testing print("word_add")
    the_word <- strsplit(word_add, " ")[[1]]
    # testing print("the_word")
    n <- length(the_word)
    # testing print(n)
    
    ########### check Bigram
    if (n == 1) {the_word <- as.character(tail(the_word,1)); functionBigram(the_word)}
    
    ################ check trigram
    else if (n == 2) {the_word <- as.character(tail(the_word,2)); functionTrigram(the_word)}
    
    ############### check quadgram
    else if (n >= 3) {the_word <- as.character(tail(the_word,3)); functionQuadgram(the_word)}
}
```



## Functioning of the App

1. The first step is to load the necessary libraries, which provide the tools and functions required for processing text data and making predictions efficiently.  
2. The app then loads the precomputed frequency matrices for bigrams, trigrams, and quadgrams. These matrices contain word pair, triplet, and quadruplet frequencies, forming the foundation of the prediction algorithm.  
3. Using this data, the app employs a prediction function to analyze input text and predict the most likely next word. This function utilizes n-gram frequency matrices to ensure accurate predictions.  
4. Finally, the core mechanism of the app is the `predictWord` function, which is optimized for fast and reliable word predictions, ensuring a seamless user experience.  

## Experience Using the App

1. The app efficiently predicts the next word by employing a well-designed prediction algorithm that leverages n-gram models. It processes the input text to deliver precise and contextually relevant results.  
2. The results are generated using robust prediction functions, ensuring accuracy and consistency. The app is intuitive and user-friendly, offering a smooth and reliable experience.  